---
title: "Predictive Analytics of Brain Stroke Vulnerability via Lifestyle and Health Factors Analysis"
author: "Anand Raj"
date: "`r Sys.Date()`"
output:
  html_document:
    code_folding: hide
    number_sections: false
    toc: yes
    toc_depth: 3
    toc_float: yes
  pdf_document:
    toc: yes
    toc_depth: '3'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Importing libraries

```{r Importing Needed Libraries, include=TRUE}
library(psych)
library(ggplot2)
```

# Import & View the Data

```{r Importing Dataset & Examing the Data, include=TRUE}
BS_Data_raw = read.csv("C:\\Users\\anand\\OneDrive\\Desktop\\HW_and_Assignments_MS_DS\\Introduction_to_Data_Science\\Project_1_EDA\\full_data.csv")

head(BS_Data_raw)

str(BS_Data_raw)
```

Inference-> Gender, Hypertension, heart_disease, ever_married, work_type, Residence_type, smoking_status, stroke are categorical variables.
Columns gender, ever_married, work_type, Residence_type, smoking_status are having "character" Data type.
Needed to be converted to Numeric/Factor before modelling -> USE Label Encoding

# Summary of Data

```{r Summary of Data, include=TRUE}
summary(BS_Data_raw)
```

Age: The mean and median donot have a huge difference, therefore less likely to have outliers.
avg_glucose_level: The mean and median donot have a huge difference, therefore less likely to have outliers.
bmi: The mean and median donot have a huge difference, therefore less likely to have outliers.

# Data Details

```{r Checks on Data, include=TRUE}
unique(BS_Data_raw$stroke)

table(BS_Data_raw$stroke)

describe(BS_Data_raw)

table(BS_Data_raw$gender)
```

## Data Preprocessing
# Checking NA Values
```{r Check for Number of NA Values in Each Column, include=TRUE}
sapply(BS_Data_raw, function(x) sum(is.na(x)))
```

Inference-> We donot have any NA/Missing values in any of the Column

# Check for Outliers
```{r Check for Outliers, include=TRUE}
ggplot(BS_Data_raw, aes(y=age)) + geom_boxplot() + labs(y="Age",title = "Age Boxplot")
```

```{r Just for Checking}
```

```{r Check2}
```