---
title: "Predictive Analytics of Brain Stroke Vulnerability via Lifestyle and Health Factors Analysis"
author: "Anand Raj"
date: "`r Sys.Date()`"
output:
  html_document:
    code_folding: hide
    number_sections: false
    toc: yes
    toc_depth: 3
    toc_float: yes
  pdf_document:
    toc: yes
    toc_depth: '3'
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r Importing the data set, include=F} 
stroke = read.csv("full_data.csv")
print(head(stroke))
```

**Importing and Summary of the Stroke Data**

```{r}
summary(stroke)
```

We are choosing on the numerical variable age, average glucose level and BMI.

Age: The mean and median do not have a huge difference, therefore less likely to have outliers.
Avg_glucose_level: The mean and median have a significant difference and, therefore likely to have outliers.
BMI: The mean and median do not have a huge difference, therefore less likely to have outliers.

**Checking for missing values in the data**

```{r Checking for missing values in the data, include=F}
missing_values <- is.na(stroke)
summary(missing_values)
```

There are no missing values in the data

**Changing the work_type children to unemployed**

```{r Changing the work_type children to unemployed,include=F}
stroke$work_type <- ifelse(stroke$work_type == "children", "unemployed", stroke$work_type)
print(head(stroke))
```

**Changing the all categorical variable to numeric variables**

```{r}
#Changing gender to numerical
stroke$gender <- as.numeric(factor(stroke$gender))

#Changing work type to numerical
stroke$work_type <- as.numeric(factor(stroke$work_type))

#Changing the location to numerical
stroke$Residence_type <- as.numeric(factor(stroke$Residence_type))

#Changing the marriage status to numerical
stroke$ever_married <- as.numeric(factor(stroke$ever_married))

#Changing the smoking status to numerical
stroke$smoking_status <- as.numeric(factor(stroke$smoking_status))

final <- stroke
head(final)
tail(final)
```

**Introducing the Age Categories**

I have grouped age into four categories such as Children, Student, Adult and Senior Citizen

```{r}
age_breaks <- c(0, 10, 18, 60, Inf)
age_labels <- c("Children", "Student", "Adult", "Senior Citizen")

# Create a new column with age categories
stroke$Category <- cut(stroke$age, breaks = age_breaks, labels = age_labels)

# Print the data set
print(head(stroke))

```

**Finding outliers of BMI and average glucose level in the data**

```{r}
bmi_outliers <- ezids::outlierKD2(stroke,bmi, rm=TRUE, boxplt=FALSE, histogram=FALSE, qqplt=TRUE)
avg_glucose_outliers <- ezids::outlierKD2(stroke,avg_glucose_level, rm=TRUE, boxplt=FALSE, histogram=FALSE, qqplt=TRUE) 
```

We have found 43 outliers in BMI and 602 in average glucose level.

**Histogram for distribution for Age**

**From this graph we understand that there are more people from age group 40-60**
```{r}
library(ggplot2)
ggplot(stroke, aes(x = age)) +
  geom_histogram(binwidth = 5, fill = "blue", color = "black", alpha = 0.7) +
  ggtitle("Use theme(plot.title = element_text(hjust = 0.5)) to center") +
  theme(plot.title = element_text(hjust = 0.5)) +
  labs(title = "Distribution of Age",
       x = "Age",
       y = "Frequency")
```

The distribution is having thick tail on both the sides.

**Histogram for distribution for Average Glucose level**

```{r}
ggplot(stroke, aes(x = avg_glucose_level)) +
  geom_histogram(binwidth = 5, fill = "lightgreen", color = "black", alpha = 0.7) +
  ggtitle("Use theme(plot.title = element_text(hjust = 0.5)) to center") +
  theme(plot.title = element_text(hjust = 0.5)) +
  labs(title = "Distribution of Average glucose level",
       x = "Average glucose level",
       y = "Frequency")
```

The graph is right skewed then the mean is greater than median. The distribution is not symmetrical. Most data points are concentrated on left.

**Distribution of BMI**

```{r}
ggplot(stroke, aes(x = bmi)) +
  geom_histogram(binwidth = 5, fill = "lightblue", color = "black", alpha = 0.7) +
  ggtitle("Use theme(plot.title = element_text(hjust = 0.5))") +
  theme(plot.title = element_text(hjust = 0.5)) +
  labs(title = "Distribution of BMI",
       x = "BMI",
       y = "Frequency")
```

The graph is almost normally distributed then the mean and median will almost be equal. The graph is also be symmetrical.

**The effect of average glucose level in hypertension**

```{r}
library(ggplot2)
stroke$gender_label <- factor(ifelse(stroke$gender == 1, "male", "female"), levels = c("male", "female"))
ggplot(stroke, aes(y = gender_label, x = avg_glucose_level, color = hypertension)) +
  geom_point() + 
  ggtitle("Use theme(plot.title = element_text(hjust = 0.5)) to center") +
  theme(plot.title = element_text(hjust = 0.5)) +
  labs(title = "Scatterplot of Average Glucose Level and Gender",y = "Gender", x = "Avg_glucose_level")
```
From this graph we understand that the females has more average glucose level compared to male

```{r}
library(ggplot2)
ggplot(stroke, aes(x = gender, y = bmi,color = gender_label)) +
  geom_boxplot() +
  ggtitle("Use theme(plot.title = element_text(hjust = 0.5)) to center") +
  theme(plot.title = element_text(hjust = 0.5)) +
  labs(
    title = "Box Plot of BMI by Gender",
    x = "Gender",
    y = "BMI"
  )
```

From this graph we understand that the average BMI Count of both gender is under 35 approximately


```{r}
library(ggplot2)
ggplot(stroke, aes(x = gender, y = avg_glucose_level,color = gender_label)) +
  geom_boxplot() +
  ggtitle("Use theme(plot.title = element_text(hjust = 0.5)) to center") +
  theme(plot.title = element_text(hjust = 0.5)) +
  labs(
    title = "Box Plot of Average Glocose Level by Gender",
    x = "Gender",
    y = "Average Glocose Level"
  )
```

From this graph we understand that the average glucose level is right skewed.

**What are the age-specific incidence and mortality rates of brain stroke?**

Null Hypothesis(H0): Age and stroke variables are independent.

Alternative Hypothesis(Ha): Age and stroke are not independent.

```{r Chi-Square Test to determine if brain stroke and age group are independent, include=TRUE}
Contingency_Table_Age <- table(stroke$Category , stroke$stroke)
Chi_Square_Test_Age_BS <- chisq.test(Contingency_Table_Age)
print(Chi_Square_Test_Age_BS$p.value)
```

The p-value for this test is <0.05 (Standard Significane level), therefore we reject Null Hypothesis. 
Hence we can conclude that Age and stroke are not independent.
Age group significantly matters in chances of having a brain stroke.

**Which category in age has highest number of Brain Strokes**

```{r Highest brain strokes in which category, include=TRUE}
print(Contingency_Table_Age)

```

```{r}
brain_stroke_yes <- subset(stroke,stroke == 1)
brain_stroke_no <- subset(stroke,stroke == 0)
```

```{r Visualization for highest number of deaths in age category, include=TRUE}
ggplot(data = brain_stroke_yes, aes(x = `Category`, fill = `Category`)) +
  geom_bar(position = "stack",col="black", fill="Orange") +
  labs(title = "Brain Strokes by Age Category",
       x = "Age Category",
       y = "Count of Brain Strokes") +
  theme(plot.title = element_text(hjust = 0.5))
```

```{r Percentage wise barplot for Age Category, include=TRUE}
library(dplyr)

# Calculate the total number of cases
total_cases <- nrow(brain_stroke_yes)

# Calculate the percentage of brain strokes for each age category
brain_stroke_yes_percentage <- brain_stroke_yes %>%
  group_by(Category) %>%
  summarise(Percentage = (n() / total_cases) * 100)

# Create the bar plot with percentages
ggplot(data = brain_stroke_yes_percentage, aes(x = `Category`, y = Percentage, fill = `Category`)) +
  geom_bar(stat = "identity", col = "black", width = 0.7, position = "stack") +
  labs(title = "Brain Strokes by Age Category (Percentage)",
       x = "Age Category",
       y = "Percentage of Brain Strokes") +
  scale_y_continuous(labels = scales::percent_format(scale = 1)) +  # Format y-axis as percentage
  theme(plot.title = element_text(hjust = 0.5))

```

The "Senior Citizen" age group exhibits the highest likelihood of experiencing a brain stroke, accounting for nearly 70% of cases

**What is the case fatality rate of stroke by gender, location, and work type**

Null Hypothesis(H0): Worktype and stroke variables are independent.

Alternative Hypothesis(Ha): Worktype and stroke are not independent.

```{r Chi-Square Test to determine if brain stroke and wortype are independent, include=TRUE}
Contingency_Table_worktype <- table(stroke$work_type , stroke$stroke)
Chi_Square_Test_worktype_BS <- chisq.test(Contingency_Table_worktype)
print(Chi_Square_Test_worktype_BS$p.value)
```

The p-value for this test is <0.05 (Standard Significane level), therefore we reject Null Hypothesis. 
Hence we can conclude that worktype and stroke are not independent.
Worktype significantly matters in chances of having a brain stroke.


**Which category in worktype has highest number of Brain Strokes***
```{r Visualization for highest number of deaths in worktype category, include=TRUE}
ggplot(data = brain_stroke_yes, aes(x = `work_type`, fill = `work_type`)) +
  geom_bar(position = "stack",col="black", fill="Green",show.legend = TRUE) +
  labs(title = "Brain Strokes by Worktype Category",
       x = "Work Type",
       y = "Count of Brain Strokes") +
  scale_fill_manual(values = c("1" = "blue", "2" = "green", "3" = "red", "4" = "purple"),
                    labels = c("Government Job", "Private", "Self Employed", "Children")) +
  theme(plot.title = element_text(hjust = 0.5))
```
```{r Percentage Wise barplot for Worktype, include=TRUE}

brain_stroke_yes <- brain_stroke_yes %>%
  mutate(work_type_named = case_when(
    work_type == 1 ~ "Government Job",
    work_type == 2 ~ "Private",
    work_type == 3 ~ "Self-Employed",
    work_type == 4 ~ "Children"
  ))


# Calculate the percentage of brain strokes for each age category
brain_stroke_yes_percentage <- brain_stroke_yes %>%
  group_by(work_type_named) %>%
  summarise(Percentage = (n() / total_cases) * 100)

# Create the bar plot with percentages
ggplot(data = brain_stroke_yes_percentage, aes(x = `work_type_named`, y = Percentage, fill = `work_type_named`)) +
  geom_bar(stat = "identity", col = "black", width = 0.7, position = "stack") +
  labs(title = "Brain Strokes by Worktype (Percentage)",
       x = "Worktype",
       y = "Percentage of Brain Strokes") +
  scale_y_continuous(labels = scales::percent_format(scale = 1)) +  # Format y-axis as percentage
  theme(plot.title = element_text(hjust = 0.5))

```

The highest number of Brain strokes is for the Private working people.

**Does brain stroke depend on gender type?**

Null Hypothesis(H0): Gender and stroke variables are independent.

Alternative Hypothesis(Ha): Gender and stroke are not independent.

```{r Chi-Square Test to determine if brain stroke and Worktype are independent, include=TRUE}
Contingency_Table_Gender <- table(stroke$gender , stroke$stroke)
Chi_Square_Test_Gender_BS <- chisq.test(Contingency_Table_Gender)
print(Chi_Square_Test_Gender_BS$p.value)
```

```{r Percentage wise stroke for Gender, include=TRUE}

# Calculate the percentage of brain strokes for each gender category
brain_stroke_yes_percentage <- brain_stroke_yes %>%
  group_by(gender_label) %>%
  summarise(Percentage = (n() / total_cases) * 100)

# Create the bar plot with percentages on the y-axis
ggplot(data = brain_stroke_yes_percentage, aes(x = gender_label, y = Percentage, fill = gender_label)) +
  geom_bar(stat = "identity", col = "black", width = 0.7, position = "stack") +
  labs(title = "Percentage of Strokes by Gender",
       x = "Gender",
       y = "Percentage of Strokes") +
  scale_y_continuous(labels = scales::percent_format(scale = 1)) +  # Format y-axis as percentage
  theme(plot.title = element_text(hjust = 0.5))

```

p-value is greater than 0.05 (Standard Significance level) and therefore we accept Null Hypothesis and so 
we can conclude that brain stroke doesn't depend on gender type.

**Does brain stroke depend on Location type?**

Null Hypothesis(H0): Location and stroke variables are independent.

Alternative Hypothesis(Ha): Location and stroke are not independent.

```{r Chi-Square Test to determine if brain stroke and Location are independent, include=TRUE}
Contingency_Table_Location <- table(stroke$Residence_type , stroke$stroke)
Chi_Square_Test_Location_BS <- chisq.test(Contingency_Table_Location)
print(Chi_Square_Test_Location_BS$p.value)
```

p-value is greater than 0.05 (Standard Significance level) and therefore we accept Null Hypothesis and so 
we can conclude that brain stroke doesn't depend on Location of residence.

```{r Percentage wise stroke for Location, include=TRUE}

brain_stroke_yes <- brain_stroke_yes %>%
  mutate(Residence_type = case_when(
    Residence_type == 1 ~ "Rural",
    Residence_type == 2 ~ "Urban"
  ))

# Calculate the percentage of brain strokes for each location category
brain_stroke_yes_percentage <- brain_stroke_yes %>%
  group_by(Residence_type) %>%
  summarise(Percentage = (n() / total_cases) * 100)

# Create the bar plot with percentages on the y-axis
ggplot(data = brain_stroke_yes_percentage, aes(x = Residence_type, y = Percentage, fill = Residence_type)) +
  geom_bar(stat = "identity", col = "black", width = 0.7, position = "stack") +
  labs(title = "Percentage of Strokes by Residence type",
       x = "Type of Residence",
       y = "Percentage of Strokes") +
  scale_y_continuous(labels = scales::percent_format(scale = 1)) +  # Format y-axis as percentage
  theme(plot.title = element_text(hjust = 0.5))
```

**Does brain stroke depend on ever-married?**

```{r Chi-Square Test to determine if brain stroke and ever-married are independent, include=TRUE}
# Chi-squared test of ever_married and stroke
contable_marr_str <- table(stroke$ever_married, stroke$stroke)
chitest_marr_str <- chisq.test(contable_marr_str)
chitest_marr_str
```

```{r Percentage wise stroke for ever-married}
# changing the labels of "ever_married"
brain_stroke_yes$ever_married <- factor(brain_stroke_yes$ever_married, levels = c("1", "2"), labels = c("unmarried(with stroke)", "married(with stroke)"))
brain_stroke_no$ever_married <- factor(brain_stroke_no$ever_married, levels = c("1", "2"), labels = c("unmarried(without stroke)", "married(without stroke)"))

married <- subset(stroke, ever_married==2)
unmarried <- subset(stroke, ever_married==1)
total_cases_married <- nrow(married)
total_cases_notmarried <- nrow(unmarried)

# Calculating the percentage of ever-married in group with stroke and without stroke respectively
married_str_percentage = nrow(subset(married, married$stroke==1)) / total_cases_married
unmarried_str_percentage = nrow(subset(unmarried, unmarried$stroke==1)) / total_cases_notmarried
Percentageofstr_married <- c(married_str_percentage, unmarried_str_percentage)
ever_married <- c("ever-married", "not ever-married")
married_strpercentage <- data.frame(Percentageofstr_married, ever_married)

# Creating the bar plot with percentage on the y-axis
ggplot(married_strpercentage, aes(x=ever_married, y=Percentageofstr_married, fill=ever_married)) + 
  geom_bar(stat = "identity", col = "black", width = 0.7) + 
  labs(x = "Ever Married or Not", y = "Percentage of Strokes", title = "Percentage of Strokes by Ever-married or Not")
```

```{r}
contable_marr_age <- table(stroke$ever_married, stroke$Category)
chitest_marr_age <- chisq.test(contable_marr_age)
chitest_marr_age
```

**How do hypertension and heart disease influence brain stroke?**

```{r Percentage wise stroke for hypertension and heart disease}
# changing the labels of "hypertension"
# brain_stroke_yes$hypertension <- factor(brain_stroke_yes$hypertension, levels = c("0", "1"), labels = c("no hypertension", "have hypertension"))
# brain_stroke_no$hypertension <- factor(brain_stroke_no$hypertension, levels = c("0", "1"), labels = c("no hypertension", "have hypertension"))

with_hyper <- subset(stroke, hypertension==1)
without_hyper <- subset(stroke, hypertension==0)
total_cases_withhyper <- nrow(with_hyper)
total_cases_withouthyper <- nrow(without_hyper)

# total_cases_yes <- nrow(brain_stroke_yes)
# total_cases_no <- nrow(brain_stroke_no)

# Calculating the percentage of hypertension in group with stroke and without stroke respectively
withhyper_str_percentage <- nrow(subset(with_hyper, with_hyper$stroke==1)) / total_cases_withhyper 
withouthyper_str_percentage = nrow(subset(without_hyper, without_hyper$stroke==1)) / total_cases_withouthyper
Percentageofstr_hyper <- c(withhyper_str_percentage, withouthyper_str_percentage)
hyper <- c("with hypertension", "without hypertension")
hyper_strpercentage <- data.frame(Percentageofstr_hyper, hyper)

ggplot(hyper_strpercentage, aes(x=hyper, y=Percentageofstr_hyper, fill=hyper)) + 
  geom_bar(stat = "identity", col = "black", width = 0.7) + 
  labs(x = "With/without Hypertension", y = "Percentage of Strokes", title = "Percentage of Strokes by Hypertension")


# changing the labels of "heart_disease"
# brain_stroke_yes$heart_disease <- factor(brain_stroke_yes$heart_disease, levels = c("0", "1"), labels = c("no heart disease", "have heart disease"))
# brain_stroke_no$heart_disease <- factor(brain_stroke_no$heart_disease, levels = c("0", "1"), labels = c("no heart disease", "have heart disease"))

with_heart <- subset(stroke, heart_disease==1)
without_heart <- subset(stroke, heart_disease==0)
total_cases_withheart <- nrow(with_heart)
total_cases_withoutheart <- nrow(without_heart)

# Calculating the percentage of heart disease in group with stroke and without stroke respectively
withheart_strpercentage <- nrow(subset(with_heart, with_heart$stroke==1)) / total_cases_withheart
withoutheart_strpercentage = nrow(subset(without_heart, without_heart$stroke==1)) / total_cases_withoutheart
Percentageofstr_heart <- c(withheart_strpercentage, withoutheart_strpercentage)
heart <- c("with heart disease", "without heart disease")
heart_strpercentage <- data.frame(Percentageofstr_heart, heart)

ggplot(heart_strpercentage, aes(x=heart, y=Percentageofstr_heart, fill=heart)) + 
  geom_bar(stat = "identity", col = "black", width = 0.7) + 
  labs(x = "With/without Heart Disease", y = "Percentage of Strokes", title = "Percentage of Strokes by Heart Disease")
```

```{r Chi-squared Test to determine the independence between hypertension, heart disease and brain stroke}
library("corrplot")
hyp_heart_str_cor <- cor(data.frame(stroke$hypertension, stroke$heart_disease, stroke$stroke))
corrplot(hyp_heart_str_cor)

contable_hyp_str <- table(stroke$hypertension, stroke$stroke)
chitest_hyp_str <- chisq.test(contable_hyp_str)
chitest_hyp_str
contable_heart_str <- table(stroke$heart_disease, stroke$stroke)
chitest_heart_str <- chisq.test(contable_heart_str)
chitest_heart_str
contable_hyp_heart <- table(stroke$hypertension, stroke$heart_disease)
chitest_hyp_heart <- chisq.test(contable_hyp_heart)
chitest_hyp_heart
```

**What are modifiable and non-modifiable risk factors for stroke in healthy people?**<br/>
We can take features like avg_glucose_level, bmi and smoking_status.
```{r}
#get responsible features from main dataset
healthy_features <- stroke[,8:11]
healthy_features$smoking_status <- replace(healthy_features$smoking_status, healthy_features$smoking_status==4, 3)
```
<br/>
For smoking status and stroke, we use chi square goodness of fit test as they are two categorical variables.<br/>
$H_0:$ smoking and stroke are independent features<br/>
$H_1:$ smoking and stroke are not independent
```{r}
# contingency table for smoking_status and stroke
contingency_table_smoking <- table(healthy_features$smoking_status, healthy_features$stroke)
print(contingency_table_smoking)
chi_smoke_status = chisq.test(contingency_table_smoking)
print(chi_smoke_status)
```
<br/>

Normality Check for Body Mass Index
```{r}
qqnorm(healthy_features$bmi, main="Body mass index", col='green') 
qqline(healthy_features$bmi)
```

This depicts body mass index is normally distributed

As p-value for the test is low, we reject null hypothesis. Now we plot bar for smoking_status vs stroke.
```{r}
# add smoking categories
healthy_features <- healthy_features %>%
  mutate(smoke_type = case_when(
    smoking_status == 1 ~ "Never smoked",
    smoking_status == 2 ~ "Formerly smoked",
    smoking_status == 3 ~ "Regular smoker"
  ))

# create barplot
ggplot(data = healthy_features, aes(x = `smoking_status`, fill = smoke_type)) +
  geom_bar(position = "stack",col="black") +
  labs(title = "Brain Strokes numbers for people",
       x = "Smoking Status",
       y = "Count of Brain Strokes") +
  theme(plot.title = element_text(hjust = 0.5))
```
<br/>
For BMI and stroke, we do a 2-sample test considering them as numerical features.<br/>
$H_0$: bmi and stroke are independent<br/>
$H_1$: bmi and stroke are not independent

```{r}
# two sample t-test
two_sample_bmi <- t.test(healthy_features$bmi, healthy_features$stroke)
two_sample_bmi
```
<br/>
As p-value for the test is low, we reject null hypothesis. Now we plot bar for BMI vs stroke.
```{r}
# add bmi_status column
healthy_features <- healthy_features %>%
  mutate(bmi_status = case_when(
    bmi > 25  ~ "Unhealthy",
    bmi < 18.5 ~ "Unhealthy",
  ))
# create barplot between bmi_status and stroke
ggplot(data = healthy_features, aes(x = `bmi`, fill = `bmi_status`)) +
  geom_bar(position = "stack") +
  labs(title = "Brain Strokes numbers for people",
       x = "BMI",
       y = "Count of Brain Strokes") +
  theme(plot.title = element_text(hjust = 0.5))
```
<br/>

Normality check for Average Glucose Level
```{r}
qqnorm(healthy_features$avg_glucose_level, main="Average glucose level", col='green') 
qqline(healthy_features$avg_glucose_level)
```
Distribution for average glucose level is not normally distributed.

For avg glucose level and stroke we use 2-sample t-test considering themas numercial variables.<br/>
$H_0$: agl and stroke are independent<br/>
$H_1$: agl and stroke are not independent
```{r}
# 2 sample t-test
two_sample_agl <- t.test(healthy_features$avg_glucose_level, healthy_features$stroke)
print(two_sample_agl)
```
<br/>
Plot histogram for average glucose level and stroke
```{r}
# add column for agl_status based on avg_glucose_level
healthy_features <- healthy_features %>%
  mutate(agl_status = case_when(
    avg_glucose_level > 100  ~ "Unhealthy",
    avg_glucose_level < 70 ~ "Unhealthy",
  ))

# create plot and fill by agl_status 
ggplot(data = healthy_features, aes(x = `avg_glucose_level`, fill = `agl_status`)) +
  geom_histogram(position = "stack",col="black") +
  labs(title = "Brain Strokes numbers for people",
       x = "Average glucose level",
       y = "Count of Brain Strokes") +
  theme(plot.title = element_text(hjust = 0.5))
```
<br/>
Average glucose level and stroke are normally distributed

<br/>
Now combining all features together and test healthy people with stroke. As two are categorical, we use chi square test.<br/>
$H_0$: Health and stroke are independent<br/>
$H_1$: Health and stroke are not independent
```{r}
# Create a column for health_status
healthy_features<- transform(healthy_features, health_status=ifelse(avg_glucose_level<100 & avg_glucose_level>70 & bmi<25 & bmi>18.5, 1,0))

healthy_table <- table(healthy_features$health_status, healthy_features$stroke)
chigoodness = chisq.test(healthy_table)
print(chigoodness$p.value)
```
<br/>
Due to low p-value, we reject null hypothesis. <br/>

Barplot between health feature and stroke
```{r}
#Create plot for health_status vs stroke
ggplot(data = healthy_features, aes(x = `health_status`, fill='health_status')) +
  geom_bar(position = "stack",col="black", fill=c("red","green")) +
  labs(title = "Brain Strokes numbers for people",
       x = "Health of people",
       y = "Count of Brain Strokes") +
  theme(plot.title = element_text(hjust = 0.5))

```

<br/>
**Q. What married life and gender are associated with an increased risk of stroke?**<br/>
We combine two columns to form another column. As two are categorical, we use chi square test.<br/>
$H_0:$ The variables are independent<br/>
$H_1:$ The variables are not independent
```{r}
# Create a dataframe with factors responsible for married_living 
df_married_living <- stroke[c('gender', 'ever_married', 'stroke')]

df_married_living <- df_married_living %>%
  mutate(married_living = case_when(
    gender==2 & ever_married==2 ~ 'Married male',
    gender==2 & ever_married==1 ~ 'Unmarried male',
    gender==1 & ever_married==2 ~ 'Married female',
    gender==1 & ever_married==1 ~ 'Unmarried female'
  ))
# contingency status for married_living and stroke
con_table <- table(df_married_living$married_living, df_married_living$stroke)
print(con_table)
chi_test <- chisq.test(con_table)
print(chi_test)
table(df_married_living$gender,df_married_living$ever_married)
```
<br/>
Due to low p-value, we reject $H_0$.<br/>
```{r}
# Create plot between married_living and stroke
ggplot(data = df_married_living, aes(x = `married_living`, fill=married_living)) +
  geom_bar(position = "stack",col="black") +
  labs(title = "Brain Strokes numbers for people",
       x = "Married life of people",
       y = "Count of Brain Strokes") +
  theme(plot.title = element_text(hjust = 0.5))
```

