---
title: "Summary Report: Predictive Analytics of Brain Stroke Vulnerability"
output:
  html_document:
    code_folding: hide
    number_sections: false
    toc: yes
    toc_depth: 3
    toc_float: yes
  #pdf_document:
    #toc: yes
    #toc_depth: '3'
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r Importing the data set, include=F} 
stroke = read.csv("full_data.csv")
print(head(stroke))
```

# Introduction
<br/>
In this report, we present an analysis of brain stroke vulnerability by examining lifestyle and health factors. We have used a dataset that includes various features such as age, average glucose level, BMI, gender, work type, residence type, ever-married status, smoking status, and the presence of hypertension and heart disease. 

# Dataset 

**Initial summary**<br/>
The dataset comprises approximately ten distinct features, including gender, age, hypertension, heart disease, marital status, work type, residence type, average glucose level, BMI, and smoking status. The primary target variable under investigation is the presence of a stroke. The dataset encompasses roughly 5000 observations. The target variable, along with two other features, heart disease and hypertension, is binary, with values of either 0 (absent) or 1 (present) to indicate the presence or absence of these conditions.

Furthermore, there are five categorical variables within the dataset: age, gender, marital status, work type, and residence type. The "gender" variable comprises values of "Male," "Female," and "Other." The "marital status" variable consists of "Yes" and "No" values. The "work type" variable encompasses categories such as "Children," "Government Job," "Private Job," "Self Employed," and "Never Worked." The "residence type" variable has values of "Rural" and "Urban." Lastly, the "smoking status" variable includes options like "Formerly Smoked," "Never Smoked," "Smokes," and "Unknown." Apart from this there is no missing value in the data set.
<br/>
**Limitations**<br/>
Firstly, the sample size of approximately 5000 observations is relatively modest, potentially limiting the dataset's ability to capture the full complexity of factors influencing stroke occurrence. Larger datasets are generally preferred for healthcare-related studies, particularly when studying rare events like strokes. Additionally, the dataset's imbalanced target variable, which contains very few instances of strokes compared to non-stroke cases, poses a significant challenge. Class imbalance can hinder the development of accurate predictive models, and addressing this issue is crucial for meaningful results. While the absence of missing values is a positive aspect, it doesn't entirely alleviate the concerns related to data quality. The dataset might still contain inaccuracies or inconsistencies that could impact the reliability of any analyses or models.

The choice of features in the dataset is another limitation. It's essential to ensure that all relevant factors influencing stroke occurrence are adequately represented. Furthermore, the categorical variables, like "smoking status" and "work type," might lack more detailed subcategories, potentially missing important nuances in these variables.

Data bias remains a concern, particularly if the dataset is drawn from specific regions or populations. Such bias can restrict the generalizability of findings and models to broader populations or healthcare contexts. The data collection methods used can introduce bias or limitations, especially if the data primarily originates from a specific type of healthcare facility. This can affect the dataset's representativeness and relevance to different healthcare settings. The dataset might not account for all potential confounding factors, such as family history, genetic predisposition, or specific lifestyle variables, which can significantly impact the likelihood of strokes.

# Analysis
Summary of Data
```{r}
# Importing and Summary of the Stroke Data
summary(stroke)
```

```{r Checking for missing values in the data, echo=FALSE, include=FALSE}
# Checking for missing values in the data
missing_values <- is.na(stroke)
summary(missing_values)
```

There are No Missing values in any of the columns.
```{r Changing the work_type children to unemployed, include = FALSE, echo = FALSE}
# Changing the work_type children to unemployed
stroke$work_type <- ifelse(stroke$work_type == "children", "unemployed", stroke$work_type)
head(stroke)
```

```{r,echo=FALSE,include=FALSE}
#Changing gender to numerical
stroke$gender <- as.numeric(factor(stroke$gender))

#Changing work type to numerical
stroke$work_type <- as.numeric(factor(stroke$work_type))

#Changing the location to numerical
stroke$Residence_type <- as.numeric(factor(stroke$Residence_type))

#Changing the marriage status to numerical
stroke$ever_married <- as.numeric(factor(stroke$ever_married))

#Changing the smoking status to numerical
stroke$smoking_status <- as.numeric(factor(stroke$smoking_status))

final <- stroke
head(final)
```
We started by importing the dataset and checking for missing values, which, fortunately, were not found in the data. We then made some data preprocessing steps, including changing the "work_type" category "children" to "unemployed" and converting categorical variables into numerical form. We categorized age into four groups: Children, Students, Adults, and Senior Citizens. This allowed us to analyze the impact of age groups on brain stroke vulnerability.

We have grouped age into four categories such as Children, Student, Adult and Senior Citizen and Encoded categorical variables.

```{r}
# Introducing the Age Categories
age_breaks <- c(0, 10, 18, 60, Inf)
age_labels <- c("Children", "Student", "Adult", "Senior Citizen")

# Create a new column with age categories
stroke$Category <- cut(stroke$age, breaks = age_breaks, labels = age_labels)

# Print the data set
print(head(stroke))

```
<br/>


```{r}
# Finding outliers of BMI and average glucose level in the data
bmi_outliers <- ezids::outlierKD2(stroke,bmi, rm=TRUE, boxplt=FALSE, histogram=FALSE, qqplt=TRUE)
avg_glucose_outliers <- ezids::outlierKD2(stroke,avg_glucose_level, rm=TRUE, boxplt=TRUE, histogram=TRUE, qqplt=TRUE) 
```
<br/>
We identified outliers in BMI and average glucose levels using the KD2 method. There were 43 outliers in BMI and 602 outliers in average glucose levels, suggesting potential data anomalies.
```{r, include = FALSE, echo = FALSE}
# Checking Skewness of Age, BMI and Average Glucose Level*
library(psych)
skew_age <- skew(stroke$age)
print(skew_age)
skew_glucose <- skew(stroke$avg_glucose_level)
print(skew_glucose)
skew_BMI <- skew(stroke$bmi)
print(skew_BMI)
```
<br/>
The skewness of age indicates a slight leftward skew in the data distribution, with more values concentrated on the right side of the mean, but the distribution is approximately symmetric. A skewness value of average glucose level and BMI indicates a right skewed distribution, with a longer tail on the right side of the mean. This suggests that there are more data points on the left side of the mean.

```{r, include = FALSE, echo = FALSE}
# **Calculating variance and standard deviation
sd_age <- sd(stroke$age)
print(sd_age)
sd_glucose <- sd(stroke$avg_glucose_level)
print(sd_glucose)
sd_BMI <- sd(stroke$bmi)
print(sd_BMI)

var_age <- var(stroke$age)
print(var_age)
var_glucose <- var(stroke$avg_glucose_level)
print(var_glucose)
var_BMI <- var(stroke$bmi)
print(var_BMI)
```

```{r, include = FALSE, echo = FALSE}
# Calculating kurtosis for age, average glucose level and BMI
library(moments)
kurtosis_age <- kurtosis(stroke$age)
print(kurtosis_age)

kurtosis_glucose <- kurtosis(stroke$avg_glucose_level)
print(kurtosis_glucose)

kurtosis_BMI <- kurtosis(stroke$bmi)
print(kurtosis_BMI)
```
<br/>
We calculated various statistical measures like mean, median, standard deviation, variance, skewness, kurtosis, and mode for the variables age, average glucose level, and BMI. These statistics provide insights into the distributions and characteristics of the data. The standard deviation value suggests that the data for average glucose levels has a wider spread than that of age and BMI. A higher variance implies that the data points deviate more from their respective means, with glucose displaying a greater spread from the mean in contrast to age and BMI. The kurtosis values for age, average glucose level, and BMI show us the shapes of their distributions. A kurtosis value of 2.004955 for age means that the distribution is close to normal, with balanced tails and a moderate peak. A kurtosis value of 4.74971 for average glucose level suggests a leptokurtic distribution, which means that it has a pronounced central peak and heavier tails, which could mean that there are outliers in the data. A kurtosis value of 2.860614 for BMI points to a mesokurtic distribution, which means that it is similar to a normal curve with moderate tail behavior.
<br/>
```{r, include = FALSE, echo = FALSE}
# **Calculation of Mode for age, average glucose level and BMI
library(DescTools)
mode_age <- Mode(stroke$age)
print(mode_age)

mode_glucose <- Mode(stroke$avg_glucose_level)
print(mode_glucose)

mode_BMI <- Mode(stroke$bmi)
print(mode_BMI)
```
Age: The modes of 78 and 102 suggest that there are two distinct age groups or that these specific ages are common in the dataset.
Average glucose level: The mode of 93.88 is a decimal value, which suggests that the data was collected precisely or that there are specific measurement practices in place. There is also another mode at 6, which suggests that there is another significant cluster of data points at this level.
BMI: The modes of 28.7 and 42 may correspond to common BMI values, which suggests that there are two categories or clusters of individuals with similar BMI values.
<br/>

<br/>
We created histograms to visualize the distributions of age, average glucose level, and BMI. These visualizations helped us understand the shape and characteristics of the data.
```{r, include=FALSE,echo=FALSE}
library(ggplot2)
```

```{r}
# **Histogram for distribution for Age**
ggplot(stroke, aes(x = age)) +
  geom_histogram(binwidth = 5, fill = "blue", color = "black", alpha = 0.7) +
  ggtitle("Use theme(plot.title = element_text(hjust = 0.5)) to center") +
  theme(plot.title = element_text(hjust = 0.5)) +
  labs(title = "Distribution of Age",
       x = "Age",
       y = "Frequency")
```
<br/>
The distribution is having thick tail on both the sides.

```{r}
ggplot(stroke, aes(x = avg_glucose_level)) +
  geom_histogram(binwidth = 5, fill = "lightgreen", color = "black", alpha = 0.7) +
  ggtitle("Use theme(plot.title = element_text(hjust = 0.5)) to center") +
  theme(plot.title = element_text(hjust = 0.5)) +
  labs(title = "Distribution of Average glucose level",
       x = "Average glucose level",
       y = "Frequency")
```
<br/>
The graph is right skewed then the mean is greater than median. The distribution is not symmetrical. Most data points are concentrated on left.
<br/>
```{r}
ggplot(stroke, aes(x = bmi)) +
  geom_histogram(binwidth = 5, fill = "lightblue", color = "black", alpha = 0.7) +
  ggtitle("Use theme(plot.title = element_text(hjust = 0.5))") +
  theme(plot.title = element_text(hjust = 0.5)) +
  labs(title = "Distribution of BMI",
       x = "BMI",
       y = "Frequency")
```
<br/>The graph is almost normally distributed then the mean and median will almost be equal. The graph is also be symmetrical.
<br/>
               **Effect of average glucose level in hypertension**
```{r}
library(ggplot2)
stroke$gender_label <- factor(ifelse(stroke$gender == 1, "male", "female"), levels = c("male", "female"))
ggplot(stroke, aes(y = gender_label, x = avg_glucose_level, color = hypertension)) +
  geom_point() + 
  ggtitle("Use theme(plot.title = element_text(hjust = 0.5)) to center") +
  theme(plot.title = element_text(hjust = 0.5)) +
  labs(title = "Scatterplot of Average Glucose Level and Gender",y = "Gender", x = "Avg_glucose_level")
```
<br/>From this graph we understand that the females has more average glucose level compared to male
<br/>
```{r}
library(ggplot2)
ggplot(stroke, aes(x = gender, y = bmi,color = gender_label)) +
  geom_boxplot() +
  ggtitle("Use theme(plot.title = element_text(hjust = 0.5)) to center") +
  theme(plot.title = element_text(hjust = 0.5)) +
  labs(
    title = "Box Plot of BMI by Gender",
    x = "Gender",
    y = "BMI"
  )
```
<br/>
From this graph we understand that the average BMI Count of both gender is under 35 approximately


```{r}
library(ggplot2)
ggplot(stroke, aes(x = gender, y = avg_glucose_level,color = gender_label)) +
  geom_boxplot() +
  ggtitle("Use theme(plot.title = element_text(hjust = 0.5)) to center") +
  theme(plot.title = element_text(hjust = 0.5)) +
  labs(
    title = "Box Plot of Average Glocose Level by Gender",
    x = "Gender",
    y = "Average Glocose Level"
  )
```
<br/>
From this graph we understand that the average glucose level is right skewed.

**What are the age-specific incidence and mortality rates of brain stroke?**
We conducted a chi-squared test to determine if age and brain stroke are independent. The results revealed that age and brain stroke are not independent. Specifically, the "Senior Citizen" age group had the highest number of brain strokes.

Null Hypothesis(H0): Age and stroke variables are independent.

Alternative Hypothesis(Ha): Age and stroke are not independent.

```{r Chi-Square Test to determine if brain stroke and age group are independent, include = FALSE, echo = FALSE}
Contingency_Table_Age <- table(stroke$Category , stroke$stroke)
Chi_Square_Test_Age_BS <- chisq.test(Contingency_Table_Age)
print(Chi_Square_Test_Age_BS$p.value)
```
<br/>
The p-value for this test is `r Chi_Square_Test_Age_BS$p.value` <0.05 (Standard Significane level), therefore we reject Null Hypothesis. 
Hence we can conclude that Age and stroke are not independent.
Age group significantly matters in chances of having a brain stroke.

**Which category in age has highest number of Brain Strokes**

```{r Highest brain strokes in which category, include=TRUE, include = FALSE, echo = FALSE}
print(Contingency_Table_Age)

```

```{r, include = FALSE, echo = FALSE}
brain_stroke_yes <- subset(stroke,stroke == 1)
brain_stroke_no <- subset(stroke,stroke == 0)
```

```{r Visualization for highest number of deaths in age category, include=TRUE}
ggplot(data = brain_stroke_yes, aes(x = `Category`, fill = `Category`)) +
  geom_bar(position = "stack",col="black", fill="Orange") +
  labs(title = "Brain Strokes by Age Category",
       x = "Age Category",
       y = "Count of Brain Strokes") +
  theme(plot.title = element_text(hjust = 0.5))
```

```{r ,include=FALSE,echo=FALSE}
library(dplyr)
```

```{r Percentage wise barplot for Age Category, include=TRUE}

# Calculate the total number of cases
total_cases <- nrow(brain_stroke_yes)

# Calculate the percentage of brain strokes for each age category
brain_stroke_yes_percentage <- brain_stroke_yes %>%
  group_by(Category) %>%
  summarise(Percentage = (n() / total_cases) * 100)

# Create the bar plot with percentages
ggplot(data = brain_stroke_yes_percentage, aes(x = `Category`, y = Percentage, fill = `Category`)) +
  geom_bar(stat = "identity", col = "black", width = 0.7, position = "stack") +
  labs(title = "Brain Strokes by Age Category (Percentage)",
       x = "Age Category",
       y = "Percentage of Brain Strokes") +
  scale_y_continuous(labels = scales::percent_format(scale = 1)) +  # Format y-axis as percentage
  theme(plot.title = element_text(hjust = 0.5))

```
<br/>
The "Senior Citizen" age group exhibits the highest likelihood of experiencing a brain stroke, accounting for nearly 70% of cases

**What is the case fatality rate of stroke by gender, location, and work type**
A chi-squared test was used to assess the relationship between work type and brain stroke. The results indicated that work type significantly matters in the likelihood of having a brain stroke, with "Private" work type having the highest number of brain strokes.
Null Hypothesis(H0): Worktype and stroke variables are independent.

Alternative Hypothesis(Ha): Worktype and stroke are not independent.

```{r Chi-Square Test to determine if brain stroke and wortype are independent, include = FALSE, echo = FALSE}
Contingency_Table_worktype <- table(stroke$work_type , stroke$stroke)
Chi_Square_Test_worktype_BS <- chisq.test(Contingency_Table_worktype)
print(Chi_Square_Test_worktype_BS$p.value)
```
<br/>
The p-value for this test is `r Chi_Square_Test_worktype_BS$p.value`<0.05 (Standard Significane level), therefore we reject Null Hypothesis. 
Hence we can conclude that worktype and stroke are not independent.
Worktype significantly matters in chances of having a brain stroke.


**Which category in worktype has highest number of Brain Strokes***
```{r Visualization for highest number of deaths in worktype category, include=TRUE}
ggplot(data = brain_stroke_yes, aes(x = `work_type`, fill = `work_type`)) +
  geom_bar(position = "stack",col="black", fill="Green",show.legend = TRUE) +
  labs(title = "Brain Strokes by Worktype Category",
       x = "Work Type",
       y = "Count of Brain Strokes") +
  scale_fill_manual(values = c("1" = "blue", "2" = "green", "3" = "red", "4" = "purple"),
                    labels = c("Government Job", "Private", "Self Employed", "Children")) +
  theme(plot.title = element_text(hjust = 0.5))
```
```{r Percentage Wise barplot for Worktype, include=TRUE}

brain_stroke_yes <- brain_stroke_yes %>%
  mutate(work_type_named = case_when(
    work_type == 1 ~ "Government Job",
    work_type == 2 ~ "Private",
    work_type == 3 ~ "Self-Employed",
    work_type == 4 ~ "Children"
  ))


# Calculate the percentage of brain strokes for each age category
brain_stroke_yes_percentage <- brain_stroke_yes %>%
  group_by(work_type_named) %>%
  summarise(Percentage = (n() / total_cases) * 100)

# Create the bar plot with percentages
ggplot(data = brain_stroke_yes_percentage, aes(x = `work_type_named`, y = Percentage, fill = `work_type_named`)) +
  geom_bar(stat = "identity", col = "black", width = 0.7, position = "stack") +
  labs(title = "Brain Strokes by Worktype (Percentage)",
       x = "Worktype",
       y = "Percentage of Brain Strokes") +
  scale_y_continuous(labels = scales::percent_format(scale = 1)) +  # Format y-axis as percentage
  theme(plot.title = element_text(hjust = 0.5))

```
<br/>
The highest number of Brain strokes is for the Private working people.

**Does brain stroke depend on gender type?**

Null Hypothesis(H0): Gender and stroke variables are independent.

Alternative Hypothesis(Ha): Gender and stroke are not independent.

```{r Chi-Square Test to determine if brain stroke and Worktype are independent, include = FALSE, echo = FALSE}
Contingency_Table_Gender <- table(stroke$gender , stroke$stroke)
Chi_Square_Test_Gender_BS <- chisq.test(Contingency_Table_Gender)
print(Chi_Square_Test_Gender_BS$p.value)
```

```{r Percentage wise stroke for Gender, include=TRUE}

# Calculate the percentage of brain strokes for each gender category
brain_stroke_yes_percentage <- brain_stroke_yes %>%
  group_by(gender_label) %>%
  summarise(Percentage = (n() / total_cases) * 100)

# Create the bar plot with percentages on the y-axis
ggplot(data = brain_stroke_yes_percentage, aes(x = gender_label, y = Percentage, fill = gender_label)) +
  geom_bar(stat = "identity", col = "black", width = 0.7, position = "stack") +
  labs(title = "Percentage of Strokes by Gender",
       x = "Gender",
       y = "Percentage of Strokes") +
  scale_y_continuous(labels = scales::percent_format(scale = 1)) +  # Format y-axis as percentage
  theme(plot.title = element_text(hjust = 0.5))

```
<br/>
p-value is greater than 0.05 (Standard Significance level) and therefore we accept Null Hypothesis and so 
we can conclude that brain stroke doesn't depend on gender type.

**Does brain stroke depend on Location type?**

Null Hypothesis(H0): Location and stroke variables are independent.

Alternative Hypothesis(Ha): Location and stroke are not independent.

```{r Chi-Square Test to determine if brain stroke and Location are independent, include = FALSE, echo = FALSE}
Contingency_Table_Location <- table(stroke$Residence_type , stroke$stroke)
Chi_Square_Test_Location_BS <- chisq.test(Contingency_Table_Location)
print(Chi_Square_Test_Location_BS$p.value)
```
p-value (`r Chi_Square_Test_Location_BS$p.value`) is greater than 0.05 (Standard Significance level) and therefore we accept Null Hypothesis and so 
we can conclude that brain stroke doesn't depend on Location of residence.

```{r Percentage wise stroke for Location, include=TRUE}

brain_stroke_yes <- brain_stroke_yes %>%
  mutate(Residence_type = case_when(
    Residence_type == 1 ~ "Rural",
    Residence_type == 2 ~ "Urban"
  ))

# Calculate the percentage of brain strokes for each location category
brain_stroke_yes_percentage <- brain_stroke_yes %>%
  group_by(Residence_type) %>%
  summarise(Percentage = (n() / total_cases) * 100)

# Create the bar plot with percentages on the y-axis
ggplot(data = brain_stroke_yes_percentage, aes(x = Residence_type, y = Percentage, fill = Residence_type)) +
  geom_bar(stat = "identity", col = "black", width = 0.7, position = "stack") +
  labs(title = "Percentage of Strokes by Residence type",
       x = "Type of Residence",
       y = "Percentage of Strokes") +
  scale_y_continuous(labels = scales::percent_format(scale = 1)) +  # Format y-axis as percentage
  theme(plot.title = element_text(hjust = 0.5))
```
<br/>
**Does brain stroke depend on ever-married?**

```{r Chi-Square Test to determine if brain stroke and ever-married are independent, include = FALSE, echo = FALSE}
# Chi-squared test of ever_married and stroke
contable_marr_str <- table(stroke$ever_married, stroke$stroke)
chitest_marr_str <- chisq.test(contable_marr_str)
chitest_marr_str
```

```{r Percentage wise stroke for ever-married}
# changing the labels of "ever_married"
brain_stroke_yes$ever_married <- factor(brain_stroke_yes$ever_married, levels = c("1", "2"), labels = c("unmarried(with stroke)", "married(with stroke)"))
brain_stroke_no$ever_married <- factor(brain_stroke_no$ever_married, levels = c("1", "2"), labels = c("unmarried(without stroke)", "married(without stroke)"))

married <- subset(stroke, ever_married==2)
unmarried <- subset(stroke, ever_married==1)
total_cases_married <- nrow(married)
total_cases_notmarried <- nrow(unmarried)

# Calculating the percentage of ever-married in group with stroke and without stroke respectively
married_str_percentage = nrow(subset(married, married$stroke==1)) / total_cases_married
unmarried_str_percentage = nrow(subset(unmarried, unmarried$stroke==1)) / total_cases_notmarried
Percentageofstr_married <- c(married_str_percentage, unmarried_str_percentage)
ever_married <- c("ever-married", "not ever-married")
married_strpercentage <- data.frame(Percentageofstr_married, ever_married)

# Creating the bar plot with percentage on the y-axis
ggplot(married_strpercentage, aes(x=ever_married, y=Percentageofstr_married, fill=ever_married)) + 
  geom_bar(stat = "identity", col = "black", width = 0.7) + 
  labs(x = "Ever Married or Not", y = "Percentage of Strokes", title = "Percentage of Strokes by Ever-married or Not")
```

```{r, include = FALSE, echo = FALSE}
contable_marr_age <- table(stroke$ever_married, stroke$Category)
chitest_marr_age <- chisq.test(contable_marr_age)
chitest_marr_age
```
<br/>
**How do hypertension and heart disease influence brain stroke?**
We examined the influence of hypertension and heart disease on brain stroke using chi-squared tests. The results showed that both hypertension and heart disease were associated with a higher likelihood of experiencing a brain stroke.
```{r Percentage wise stroke for hypertension and heart disease}
# changing the labels of "hypertension"
# brain_stroke_yes$hypertension <- factor(brain_stroke_yes$hypertension, levels = c("0", "1"), labels = c("no hypertension", "have hypertension"))
# brain_stroke_no$hypertension <- factor(brain_stroke_no$hypertension, levels = c("0", "1"), labels = c("no hypertension", "have hypertension"))

with_hyper <- subset(stroke, hypertension==1)
without_hyper <- subset(stroke, hypertension==0)
total_cases_withhyper <- nrow(with_hyper)
total_cases_withouthyper <- nrow(without_hyper)

# total_cases_yes <- nrow(brain_stroke_yes)
# total_cases_no <- nrow(brain_stroke_no)

# Calculating the percentage of hypertension in group with stroke and without stroke respectively
withhyper_str_percentage <- nrow(subset(with_hyper, with_hyper$stroke==1)) / total_cases_withhyper 
withouthyper_str_percentage = nrow(subset(without_hyper, without_hyper$stroke==1)) / total_cases_withouthyper
Percentageofstr_hyper <- c(withhyper_str_percentage, withouthyper_str_percentage)
hyper <- c("with hypertension", "without hypertension")
hyper_strpercentage <- data.frame(Percentageofstr_hyper, hyper)

ggplot(hyper_strpercentage, aes(x=hyper, y=Percentageofstr_hyper, fill=hyper)) + 
  geom_bar(stat = "identity", col = "black", width = 0.7) + 
  labs(x = "With/without Hypertension", y = "Percentage of Strokes", title = "Percentage of Strokes by Hypertension")


# changing the labels of "heart_disease"
# brain_stroke_yes$heart_disease <- factor(brain_stroke_yes$heart_disease, levels = c("0", "1"), labels = c("no heart disease", "have heart disease"))
# brain_stroke_no$heart_disease <- factor(brain_stroke_no$heart_disease, levels = c("0", "1"), labels = c("no heart disease", "have heart disease"))

with_heart <- subset(stroke, heart_disease==1)
without_heart <- subset(stroke, heart_disease==0)
total_cases_withheart <- nrow(with_heart)
total_cases_withoutheart <- nrow(without_heart)

# Calculating the percentage of heart disease in group with stroke and without stroke respectively
withheart_strpercentage <- nrow(subset(with_heart, with_heart$stroke==1)) / total_cases_withheart
withoutheart_strpercentage = nrow(subset(without_heart, without_heart$stroke==1)) / total_cases_withoutheart
Percentageofstr_heart <- c(withheart_strpercentage, withoutheart_strpercentage)
heart <- c("with heart disease", "without heart disease")
heart_strpercentage <- data.frame(Percentageofstr_heart, heart)

ggplot(heart_strpercentage, aes(x=heart, y=Percentageofstr_heart, fill=heart)) + 
  geom_bar(stat = "identity", col = "black", width = 0.7) + 
  labs(x = "With/without Heart Disease", y = "Percentage of Strokes", title = "Percentage of Strokes by Heart Disease")
```

```{r Chi-squared Test to determine the independence between hypertension, heart disease and brain stroke, include = FALSE, echo = FALSE}
library("corrplot")
hyp_heart_str_cor <- cor(data.frame(stroke$hypertension, stroke$heart_disease, stroke$stroke))
corrplot(hyp_heart_str_cor)

contable_hyp_str <- table(stroke$hypertension, stroke$stroke)
chitest_hyp_str <- chisq.test(contable_hyp_str)
chitest_hyp_str
contable_heart_str <- table(stroke$heart_disease, stroke$stroke)
chitest_heart_str <- chisq.test(contable_heart_str)
chitest_heart_str
contable_hyp_heart <- table(stroke$hypertension, stroke$heart_disease)
chitest_hyp_heart <- chisq.test(contable_hyp_heart)
chitest_hyp_heart
```
<br/>
**What are modifiable and non-modifiable risk factors for stroke in healthy people?**<br/>
We analyzed modifiable and non-modifiable risk factors for stroke in healthy individuals by considering average glucose level, BMI, and smoking status. We conducted hypothesis tests and visualizations to understand their relationships with stroke risk. We can take features like avg_glucose_level, bmi and smoking_status.
```{r, include = FALSE, echo = FALSE}
#get responsible features from main dataset
healthy_features <- stroke[,8:11]
healthy_features$smoking_status <- replace(healthy_features$smoking_status, healthy_features$smoking_status==4, 3)
```
<br/>
For smoking status and stroke, we use chi square goodness of fit test as they are two categorical variables.<br/>
$H_0:$ smoking and stroke are independent features<br/>
$H_1:$ smoking and stroke are not independent
```{r}
# contingency table for smoking_status and stroke
contingency_table_smoking <- table(healthy_features$smoking_status, healthy_features$stroke)
#print(contingency_table_smoking)
chi_smoke_status = chisq.test(contingency_table_smoking)
print(chi_smoke_status)
```
<br/>

Normality Check for Body Mass Index
```{r}
qqnorm(healthy_features$bmi, main="Body mass index", col='green') 
qqline(healthy_features$bmi)
```
<br/>
This depicts body mass index is normally distributed

As p-value for the test is low, we reject null hypothesis. Now we plot bar for smoking_status vs stroke.
```{r}
# add smoking categories
healthy_features <- healthy_features %>%
  mutate(smoke_type = case_when(
    smoking_status == 1 ~ "Never smoked",
    smoking_status == 2 ~ "Formerly smoked",
    smoking_status == 3 ~ "Regular smoker"
  ))

# create barplot
ggplot(data = healthy_features, aes(x = `smoking_status`, fill = smoke_type)) +
  geom_bar(position = "stack",col="black") +
  labs(title = "Brain Strokes numbers for people",
       x = "Smoking Status",
       y = "Count of Brain Strokes") +
  theme(plot.title = element_text(hjust = 0.5))
```
<br/>
For BMI and stroke, we do a 2-sample test considering them as numerical features.<br/>
$H_0$: bmi and stroke are independent<br/>
$H_1$: bmi and stroke are not independent

```{r, include = FALSE, echo = FALSE}
# two sample t-test
two_sample_bmi <- t.test(healthy_features$bmi, healthy_features$stroke)
two_sample_bmi
```
<br/>
As p-value (`r two_sample_bmi$p.value`) for the test is low, we reject null hypothesis. Now we plot bar for BMI vs stroke.
```{r}
# add bmi_status column
healthy_features <- healthy_features %>%
  mutate(bmi_status = case_when(
    bmi > 25  ~ "Unhealthy",
    bmi < 18.5 ~ "Unhealthy",
  ))
# create barplot between bmi_status and stroke
ggplot(data = healthy_features, aes(x = `bmi`, fill = `bmi_status`)) +
  geom_bar(position = "stack") +
  labs(title = "Brain Strokes numbers for people",
       x = "BMI",
       y = "Count of Brain Strokes") +
  theme(plot.title = element_text(hjust = 0.5))
```
<br/>

Normality check for Average Glucose Level
```{r}
qqnorm(healthy_features$avg_glucose_level, main="Average glucose level", col='green') 
qqline(healthy_features$avg_glucose_level)
```
<br/>Distribution for average glucose level is not normally distributed.

For avg glucose level and stroke we use 2-sample t-test considering themas numercial variables.<br/>
$H_0$: agl and stroke are independent<br/>
$H_1$: agl and stroke are not independent
```{r, include = FALSE, echo = FALSE}
# 2 sample t-test
two_sample_agl <- t.test(healthy_features$avg_glucose_level, healthy_features$stroke)
print(two_sample_agl)
```
<br/>
Plot histogram for average glucose level and stroke
```{r}
# add column for agl_status based on avg_glucose_level
healthy_features <- healthy_features %>%
  mutate(agl_status = case_when(
    avg_glucose_level > 100  ~ "Unhealthy",
    avg_glucose_level < 70 ~ "Unhealthy",
  ))

# create plot and fill by agl_status 
ggplot(data = healthy_features, aes(x = `avg_glucose_level`, fill = `agl_status`)) +
  geom_histogram(position = "stack",col="black") +
  labs(title = "Brain Strokes numbers for people",
       x = "Average glucose level",
       y = "Count of Brain Strokes") +
  theme(plot.title = element_text(hjust = 0.5))
```
<br/>
Average glucose level and stroke are normally distributed

Now combining all features together and test healthy people with stroke. As two are categorical, we use chi square test.<br/>
$H_0$: Health and stroke are independent<br/>
$H_1$: Health and stroke are not independent
```{r, include = FALSE, echo = FALSE}
# Create a column for health_status
healthy_features<- transform(healthy_features, health_status=ifelse(avg_glucose_level<100 & avg_glucose_level>70 & bmi<25 & bmi>18.5, 1,0))

healthy_table <- table(healthy_features$health_status, healthy_features$stroke)
chigoodness = chisq.test(healthy_table)
print(chigoodness$p.value)
```
<br/>
Due to low p-value of `r chigoodness$p.value`, we reject null hypothesis. <br/>

Barplot between health feature and stroke
```{r}
#Create plot for health_status vs stroke
ggplot(data = healthy_features, aes(x = `health_status`, fill='health_status')) +
  geom_bar(position = "stack",col="black", fill=c("red","green")) +
  labs(title = "Brain Strokes numbers for people",
       x = "Health of people",
       y = "Count of Brain Strokes") +
  theme(plot.title = element_text(hjust = 0.5))

```

<br/>
**What married life and gender are associated with an increased risk of stroke?**<br/>
We combine two columns to form another column. As two are categorical, we use chi square test.<br/>
$H_0:$ The variables are independent<br/>
$H_1:$ The variables are not independent
```{r, include = FALSE, echo = FALSE}
# Create a dataframe with factors responsible for married_living 
df_married_living <- stroke[c('gender', 'ever_married', 'stroke')]

df_married_living <- df_married_living %>%
  mutate(married_living = case_when(
    gender==2 & ever_married==2 ~ 'Married male',
    gender==2 & ever_married==1 ~ 'Unmarried male',
    gender==1 & ever_married==2 ~ 'Married female',
    gender==1 & ever_married==1 ~ 'Unmarried female'
  ))
# contingency status for married_living and stroke
con_table <- table(df_married_living$married_living, df_married_living$stroke)
print(con_table)
chi_test <- chisq.test(con_table)
print(chi_test)
table(df_married_living$gender,df_married_living$ever_married)
```
<br/>
Due to low p-value of `r chi_test$p.value`, we reject $H_0$.<br/>
```{r}
# Create plot between married_living and stroke
ggplot(data = df_married_living, aes(x = `married_living`, fill=married_living)) +
  geom_bar(position = "stack",col="black") +
  labs(title = "Brain Strokes numbers for people",
       x = "Married life of people",
       y = "Count of Brain Strokes") +
  theme(plot.title = element_text(hjust = 0.5))
```

# Conclusion
This analysis provides valuable insights into factors that influence brain stroke vulnerability. Age, hypertension, heart disease, average glucose levels, and BMI are significant factors associated with an increased risk of brain stroke. Additionally, "Married males" are more prone to stroke. This information can be used to develop preventive measures and interventions to reduce the risk of stroke in susceptible populations.

# Recommendations
Based on the analysis, we recommend the following:

* Encourage regular health check-ups, especially for individuals with higher average glucose levels.
* Promote a healthy lifestyle, including maintaining a healthy BMI.
* Provide targeted awareness campaigns about the risks of stroke among specific age groups.
* Consider tailored interventions for married males to reduce their stroke risk.
* By addressing these recommendations, we can potentially reduce the incidence of strokes and improve public health.
